### 2022-06-21 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f716db62355123121d22ac5466a04778c04fea61 🚀

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that reservoir richness
alone, although a component of viral hazard, is not a sufficiently
integrative predictor of risk. We offer alternative insights based on
viral sharing, host compositional uniqueness, and host phylogenetic
diversity and phylogeographic regions. By comparing our aggregated
measure of risk to a proxy for human density, namely the proportion of
each pixel that is covered by urban or built land, we provide a
synthetic risk map, allowing the identification of hotspots where the
bat-betacoronavirus system may originate novel viruses in humans.

@@

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately track host diversity, with some
areas having over-distinct viruses (eastern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions highlight that southeast Asia is a likely hostpot of
unconfirmed hosts of betacoronaviruses,\textsuperscript{21} which would
likely result in additional viral discoveries. This idea is unsurprising
given the growing realization, especially since the emergence of
SARS-CoV-2, that unique lineages of similar viruses are widespread in
bats but still mostly undescribed. The most distinct
bats-betacoronaviruses complex is found in South America, a region with
a comparatively lower number of hosts; this matches with the isolation
through viariance of the host group, and may highlight a different
co-evolutionary dynamic. Alternatively, this distinctiveness hostpot may
be a product of under-sampling: South-America is one of the places where
the fewest \emph{Betacoronavirus} sequences have been
discovered,\textsuperscript{2,22,23} resulting in sparser phylogenetic
tree, thereby artificially inflating distinctiveness. Adding more
viruses would bring the distinctiveness of known sequences down.
Finally, South-America is the range of Phyllostomidae, a group of bats
that underwent explosive diversification events,\textsuperscript{20}
which may drive the emergence of multiple viral lineages.

\hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
\subsection{The phylogeographic regions of hosts and their viruses
overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}

@@

{[}Tim tinker with para about conservation{]} There are several factors
that drive changes in the diversity of bats,\textsuperscript{51} but
human activities' effects on the ecosystem (like modifications of land
use) could significantly decrease it. Therefore, it can be suggested
that changes in the diversity of betacoronaviruses in bats are linked to
their biogeographic variation, and human population density and other
anthropogenic factors are decisive moderators for its implications in
public health. With the increase of contact between humans and potential
hosts, we also increase the risk of emergence of novel
diseases,\textsuperscript{52} as previous studies on RNA viruses suggest
the importance of host phylogeography at the time of virus
dispersal.\textsuperscript{53} One of these scenarios where interaction
between bats and humans can occur can be seed dispersal in tropical
agroecosystems. It opens the discussion of whether the fruits thrown by
bats not only disperse seeds but could also be a source of indirect
interaction between viruses of bat origin and
humans.\textsuperscript{54} This represents a challenge for conservation
strategies and disease ecology since some areas can haveboth potential
for the acquisition of zoonotic viruses and bat-human interactions; in
particular, the challenge lies in the fact that actual exposure must
then be quantified accounting for several transmission scenarios,
including both direct and indirect bat - human interaction.

\textbf{Acknowledgements}: We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. This work was supported by funding to the Viral Emergence
Research Initiative (VERENA) consortium including NSF BII 2021909 and a
grant from Institut de Valorisation des Données (IVADO). This research
was enabled in part by support provided by Calcul Québec
(www.calculquebec.ca) and Compute Canada (www.computecanada.ca). NF is
funded by the NSERC BIOS² CREATE program. TP and NF are funded by the
Courtois Foundation. RLM was supported by Bryce Carmine and Anne Carmine
(née Percival), through the Massey University Foundation.

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{55} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{56} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{57} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{58} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{59} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{60} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{61} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range
overlap.\textsuperscript{62} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{63} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{64,65} in that this allows
multiple independent host-virus coevolutionary dynamics to take place in
the same location. In the colorimetric space, this correspond to yellow
-- because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{see \emph{e.g.} 66} we measured the risk
level by calculating the angular distance of the hue of each pixel to a
reference value of 60, and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
population trends reveal key predictors of virus spillover risk.
\emph{Proceedings of the Royal Society B: Biological Sciences}
\textbf{287}, 20192736 (2020).}

\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
The Importance of Host Phylogeographic Structure in the Spatial Spread
of Arenaviruses. \emph{PLOS Pathogens} \textbf{13}, e1006073 (2017).}

\leavevmode\hypertarget{ref-Deshpande2022ForFru}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Deshpande, K., Vanak, A. T., Devy, M. S. \&
Krishnaswamy, J. Forbidden fruits? Ecosystem services from seed
dispersal by fruit bats in the context of latent zoonotic risk.
\emph{Oikos (Copenhagen, Denmark)} oik.08359 (2022)
doi:\href{https://doi.org/10.1111/oik.08359}{10.1111/oik.08359}.}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{55. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{56. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
ecological uniqueness over broad spatial extents using species
distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
\CSLLeftMargin{62. }
\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
Predicting the global mammalian viral sharing network using
phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}

\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
\CSLLeftMargin{63. }
\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
geography of lake carbon cycling. \emph{Limnology and Oceanography
Letters} \textbf{3}, 49--56 (2018).}

\leavevmode\hypertarget{ref-Cavender-Bares2009MerCom}{}%
\CSLLeftMargin{64. }
\CSLRightInline{Cavender-Bares, J., Kozak, K. H., Fine, P. V. A. \&
Kembel, S. W. The merging of community ecology and phylogenetic biology.
\emph{Ecol. Lett.} \textbf{12}, 693--715 (2009).}

>>>>

\vfill
Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.

@@

It is noteworthy that outside of South America, viral evolutionary
distinctiveness does not accurately track host diversity, with some
areas having over-distinct viruses (eastern China but, oddly, not the
rest of southeast Asia). There are a number of likely explanations.
First, given the richness of bats in southeast Asia, many
betacoronaviruses likely remain to be discovered in this region. Indeed,
global predictions highlight that southeast Asia is a likely hostpot of
unconfirmed hosts of betacoronaviruses,\textsuperscript{21} which would
likely result in additional viral discoveries. This idea is unsurprising
given the growing realization, especially since the emergence of
SARS-CoV-2, that unique lineages of similar viruses are widespread in
bats but still mostly undescribed. The most distinct
bats-betacoronaviruses complex is found in South America, a region with
a comparatively lower number of hosts; this matches with isolation
through host vicariance, and may highlight a different co-evolutionary
dynamic. Alternatively, this distinctiveness hostpot may be a product of
under-sampling: South-America is one of the places where the fewest
\emph{Betacoronavirus} sequences have been
discovered,\textsuperscript{2,22,23} resulting in sparser phylogenetic
tree, thereby artificially inflating distinctiveness. Adding more
viruses would bring the distinctiveness of known sequences down.
Finally, South-America is the range of Phyllostomidae, a group of bats
that underwent explosive diversification events,\textsuperscript{20}
which may drive the emergence of multiple viral lineages.

\hypertarget{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}{%
\section{The phylogeographic regions of hosts and their viruses
overlap}\label{the-phylogeographic-regions-of-hosts-and-their-viruses-overlap}}

@@

The diversity and diversification potential of bats responds to
anthropogenic factors others than shifting climates.\textsuperscript{51}
Land use changes could significantly decrease bat suitability, notably
through effects on diet and availability of
habitats.\textsuperscript{52} As our results establish that the
diversification of bats betacoronaviruses happens on top of processes
affecting hosts, biogeographic variation in human population density and
anthropogenic disturbances may feed into co-evolutionary dynamics.
Increase in humans-hosts contacts also increase the risk of emergence of
novel diseases,\textsuperscript{53} so does the changes in landscape
connectivity at local/regional scales.\textsuperscript{54} This
represents a challenge for both conservation strategies and disease
ecology: some areas can a high emergence risk and more potential for the
acquisition of zoonotic viruses through bat-human
encounters.\textsuperscript{55} In particular, the challenge ahead lies
in the need to quantify actual exposure (and risk) accounting for
several transmission scenarios, including both direct and indirect bat -
human interactions, and feeding back into the provision of ecosystem
services by bats.

\textbf{Acknowledgements}: We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. This work was supported by funding to the Viral Emergence
Research Initiative (VERENA) consortium including NSF BII 2021909 and a
grant from Institut de Valorisation des Données (IVADO). This research
was enabled in part by support provided by Calcul Québec
(www.calculquebec.ca) and Compute Canada (www.computecanada.ca). NF is
funded by the NSERC BIOS² CREATE program. TP and NF are funded by the
Courtois Foundation. RLM was supported by Bryce Carmine and Anne Carmine
(née Percival), through the Massey University Foundation. DJB was
supported by the National Institute of General Medical Sciences of the
National Institutes of Health (P20GM134973).

@@

We downloaded the rangemap of every current bat species that was
classified as an empirically documented host of \emph{Betacoronavirus}
from the previous step, according to recent IUCN
data.\textsuperscript{56} The range maps were subsequently rasterized
using the \texttt{rasterize} function from
\texttt{GDAL}\textsuperscript{57} at a resolution of approximately
100kmx100km. For every pixel in the resulting raster where at least one
bat host of \emph{Betacoronavirus} was present, we extract the species
pool (list of all known bat hosts), which was used to calculate the
following risk assessment components: bat phylogenetic diversity, bat
compositional uniqueness, and predicted viral sharing risk.

@@

For every pixel, we measured Faith's Phylogenetic
Diversity\textsuperscript{58} based on a recent synthetic tree with
robust time calibration, covering about 6000 mammalian
species.\textsuperscript{59} Faith's PD measures the sum of unique
branches from an arbitrary root to a set of tips, and comparatively
larger values indicate a more phylogenetic diverse species pool. We
measured phylogenetic diversity starting from the root of the entire
tree (and not from Chiroptera); this bears no consequences on the
resulting values, since all branches leading up to Chiroptera are only
counted one per species pool, and (as we explain when describing the
assembly of the composite risk map), all individual risk components are
ranged in {[}0,1{]}. This measure incorporates a richness component,
which we chose not to correct for; the interpretation of the
phylogenetic diversity is therefore a weighted species richness, that
accounts for phylogenetic over/under-dispersal in some places.

@@

For every species pool, we measured its Local Contribution to
Beta-Diversity;\textsuperscript{60} LCBD works from a species-data
matrix (traditionally noted as \(\mathbf{Y}\)), where species are rows
and sites are columns, and a value of 1 indicates occurrence. We
extracted the Y matrix assuming that every pixel represents a unique
location, and following best practices\textsuperscript{61} transformed
it using Hellinger's distance to account for unequal bat richness at
different pixels. The correction of raw community data is particularly
important for two reasons: first, it prevents the artifact of richer
sites having higher importance; second, it removes the effect of overall
species richness, which is already incorporated in the phylogenetic
diversity component. High values of LCBD indicate that the pixel has a
community that is on average more dissimilar in species composition than
what is expected knowing the entire matrix, i.e.~a more unique
community. Recent results by\textsuperscript{62} shows that LCBD
measures are robust with regards to spatial scale, and are therefore
applicable at the global scale.

@@

For all bat hosts of \emph{Betacoronavirus}, we extracted their
predicted viral sharing network, generated from a previously published
generalized additive mixed model of virus sharing by a tensor function
of phylogenetic distance and geographic range overlap across
mammals.\textsuperscript{63} This network stores pairwise values of
viral community similarity. To project viral sharing values into a
single value for every pixel, we averaged the pairwise scores. High
values of the average sharing propensity means that this specific extant
bat assemblage is likely to be proficient at exchanging viruses.

@@

To visualize the aggregated risk at the global scale, we combine the
three individual risk components (phylogenetic diversity, compositional
uniqueness, and viral sharing) using an additive color
model.\textsuperscript{64} In this approach, every risk component gets
assigned a component in the RGB color model (phylogenetic diversity is
green, compositional uniqueness is red, and viral sharing is blue). In
order to achieve a valid RGB measure, all components are re-scaled to
the {[}0,1{]} interval, so that a pixel with no sharing, no phylogenetic
diversity, and no compositional uniqueness is black, and a pixel with
maximal values for each is white. This additive model conveys both the
intensity of the overall risk, but also the nature of the risk as colors
diverge towards combinations of values for three risk components. Out of
the possible combinations, the most risky in terms or rapid
diversification and spillover potential is high phylogenetic diversity
and low viral sharing,\textsuperscript{65} in that this allows multiple
independent host-virus coevolutionary dynamics to take place in the same
location. In the colorimetric space, this correspond to yellow --
because the HSV space is more amenable to calculations for feature
extraction,\textsuperscript{66} we measured the risk level by
calculating the angular distance of the hue of each pixel to a reference
value of 60 (yellow), and weighted this risk level by the value
component. Specifically, given a pixel with colorimetric coordinates
\((h,s,v)\), its ranged weighted risk value is

@@

\leavevmode\hypertarget{ref-Treitler2016EffLoc}{}%
\CSLLeftMargin{52. }
\CSLRightInline{Treitler, J. T., Heim, O., Tschapka, M. \& Jung, K. The
effect of local land use and loss of forests on bats and nocturnal
insects. \emph{Ecology and Evolution} \textbf{6}, 4289--4297 (2016).}

\leavevmode\hypertarget{ref-Johnson2020GloShi}{}%
\CSLLeftMargin{53. }
\CSLRightInline{Johnson, C. K. \emph{et al.} Global shifts in mammalian
population trends reveal key predictors of virus spillover risk.
\emph{Proceedings of the Royal Society B: Biological Sciences}
\textbf{287}, 20192736 (2020).}

\leavevmode\hypertarget{ref-Gryseels2017WheVir}{}%
\CSLLeftMargin{54. }
\CSLRightInline{Gryseels, S. \emph{et al.} When Viruses Don't Go Viral:
The Importance of Host Phylogeographic Structure in the Spatial Spread
of Arenaviruses. \emph{PLOS Pathogens} \textbf{13}, e1006073 (2017).}

\leavevmode\hypertarget{ref-Amman2011InvRol}{}%
\CSLLeftMargin{55. }
\CSLRightInline{Amman, B. R. \emph{et al.} \emph{Investigating the role
of bats in emerging zoonoses: Balancing ecology, conservation and public
health interest}. (FAO, 2011).}

\leavevmode\hypertarget{ref-IUCN2021IucRed}{}%
\CSLLeftMargin{56. }
\CSLRightInline{IUCN. The IUCN Red List of Threatened Species.
\url{https://www.iucnredlist.org} (2021).}

\leavevmode\hypertarget{ref-RouaultEven2022GdaOgr}{}%
\CSLLeftMargin{57. }
\CSLRightInline{Rouault, E. \emph{et al.} \emph{GDAL/OGR Geospatial Data
Abstraction software Library}. (Zenodo, 2022).
doi:\href{https://doi.org/10.5281/ZENODO.5884351}{10.5281/ZENODO.5884351}.}

\leavevmode\hypertarget{ref-Faith1992ConEva}{}%
\CSLLeftMargin{58. }
\CSLRightInline{Faith, D. P. Conservation evaluation and phylogenetic
diversity. \emph{Biological Conservation} \textbf{61}, 1--10 (1992).}

\leavevmode\hypertarget{ref-Upham2019InfMam}{}%
\CSLLeftMargin{59. }
\CSLRightInline{Upham, N. S., Esselstyn, J. A. \& Jetz, W. Inferring the
mammal tree: Species-level sets of phylogenies for questions in ecology,
evolution, and conservation. \emph{PLOS Biology} \textbf{17}, e3000494
(2019).}

\leavevmode\hypertarget{ref-Legendre2013BetDiv}{}%
\CSLLeftMargin{60. }
\CSLRightInline{Legendre, P. \& De Cáceres, M. Beta diversity as the
variance of community data: Dissimilarity coefficients and partitioning.
\emph{Ecology Letters} \textbf{16}, 951--963 (2013).}

\leavevmode\hypertarget{ref-Legendre2019SpaTem}{}%
\CSLLeftMargin{61. }
\CSLRightInline{Legendre, P. \& Condit, R. Spatial and temporal analysis
of beta diversity in the Barro Colorado Island forest dynamics plot,
Panama. \emph{Forest Ecosystems} \textbf{6}, 7 (2019).}

\leavevmode\hypertarget{ref-Dansereau2022EvaEco}{}%
\CSLLeftMargin{62. }
\CSLRightInline{Dansereau, G., Legendre, P. \& Poisot, T. Evaluating
ecological uniqueness over broad spatial extents using species
distribution modelling. \emph{Oikos} \textbf{n/a}, e09063 (2022).}

\leavevmode\hypertarget{ref-Albery2020PreGlo}{}%
\CSLLeftMargin{63. }
\CSLRightInline{Albery, G. F., Eskew, E. A., Ross, N. \& Olival, K. J.
Predicting the global mammalian viral sharing network using
phylogeography. \emph{Nature Communications} \textbf{11}, 2260 (2020).}

\leavevmode\hypertarget{ref-Seekell2018GeoLak}{}%
\CSLLeftMargin{64. }
\CSLRightInline{Seekell, D. A., Lapierre, J.-F. \& Cheruvelil, K. S. A
geography of lake carbon cycling. \emph{Limnology and Oceanography
Letters} \textbf{3}, 49--56 (2018).}

### 2022-06-21 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@f716db62355123121d22ac5466a04778c04fea61 🚀

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that reservoir richness
alone, although a component of viral hazard, is not a sufficiently
integrative predictor of risk. We offer alternative insights based on
viral sharing, host compositional uniqueness, and host phylogenetic
diversity and phylogeographic regions. By comparing our aggregated
measure of risk to a proxy for human density, namely the proportion of
each pixel that is covered by urban or built land, we provide a
synthetic risk map, allowing the identification of hotspots where the
bat-betacoronavirus system may originate novel viruses in humans.

>>>>

Driven by the need to understand the ecological factors involved in the
emergence of \emph{Betacoronavirus} (the genus causing the SARS, MERS,
and COVID-19 in human) through bat hosts, we develop an approach to the
assesment of spillover risk based on the Geographic Mosaic Theory of
Coevolution. In doing so, we provide a global mapping of the spillover
risk posed by betacoronaviruses, reflecting the fact that this risk is
best understood through the multi-faceted prism of ecological and
evolutionary mechanisms. Our framework reveals that host richness alone,
although a component of viral hazard, is not a sufficiently integrative
predictor of risk. We offer alternative insights based on viral sharing,
host compositional uniqueness, and host phylogenetic diversity and
phylogeographic regions. By comparing our aggregated measure of risk to
a proxy for human density, namely the proportion of each pixel that is
covered by urban or built land, we provide a synthetic risk map,
allowing the identification of hotspots where the bat-betacoronavirus
interaction network may facilitate the emergence of novel viruses and
their spillover into human populations.
